Comment: Inferring Trade Costs from Trade Booms and Trade Busts

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

This is potentially an important result that sheds light on the globalization processes.
However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
The authors use a weighted arithmetic average of bilateral transportation costs to compute country-specific trade costs. Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows\footnote{namely trade flows between a pair of countries} ($\sigma > 1$ is elasticity of substitution).
In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
Because $1/(1-\sigma)$ < 0, the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
This is not compensated by the weight they use (end-of-period GDP).
Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.

In this paper, we use a three-countries version of \cite{AW2003}'s model for the aggregation of bilateral trade costs shows that there was no difference in nature between the two globalizations to explain the issue.
More generally, we cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade flows and excludes a priori other possibles causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.
Therefore, Jacks \textit{et al.}'s approach cannot be an alternative to traditional investigations of impediments to trade, such as commodity price gaps.

>>>>

This is potentially an important result that sheds light on the globalization processes.
However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
The authors use a weighted arithmetic average of bilateral transportation costs to compute country-specific trade costs. Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows\footnote{namely trade flows between a pair of countries} ($\sigma > 1$ is elasticity of substitution).
In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
Because $1/(1-\sigma)$ < 0, the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
This is not compensated by the weight they use (end-of-period GDP).
Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade; this is not the case in the second wave of globalization.

In this paper, we use a three-country version of \cite{AW2003}'s model for the aggregation of bilateral trade costs and show that there was no difference in nature between the two globalizations.
More generally, we cast doubts on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.
Therefore, Jacks \textit{et al.}'s approach cannot be an alternative to traditional investigations of impediments to trade, such as commodity price gaps.
